Reproducible Research: Peer Assessment 1
=============================================


## Loading and preprocessing the data

```{r}
data <-read.csv("./activity.csv",sep=",",head=T,colClasses=c("integer","Date","integer"))
head(data)
```


## What is mean total number of steps taken per day?
1. Change time locale for R to English.
2. Split data by day (date) and calculate steps in each one.
3. Plot the histogamme.
4. Calculate mean and medaine.

```{r}
Sys.setlocale("LC_TIME", "English")
totalStepsPerDay <- sapply(split(data$steps,data$date),sum,na.rm=T)

hist(totalStepsPerDay)

meanPerDay <- mean(totalStepsPerDay )
MedianPerDay <- median(totalStepsPerDay)
```
The mean of number of steps is **`r meanPerDay`** and mediane is **`r MedianPerDay`.**  

## What is the average daily activity pattern?
1. Split data by intervals.
2. Calculate average of steps in each 5 minutes interval.
3. Plot 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis).
4. Find the interval that contain maximum number of steps. 

```{r}
avrgPerInterval<-sapply(split(data$steps,data$interval),mean,na.rm=T)
intervals<- unique(data$interval)
plot(x=intervals,y=avrgPerInterval,type='l',xlab='5-minute intervals',ylab='average number of step')

totPerInterval<-sapply(split(data$steps,data$interval),sum,na.rm=T)
indMax <- which.max(totPerInterval)
intervaleMax <- intervals[indMax]

```

The 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps is **`r intervaleMax`** and it contain **`r totPerInterval[indMax]`** steps. 
## Imputing missing values

Remove all missing values form data and build new dataSet.
```{r}
nrow(na.omit(data))
NewData<-data
for(i in 1:nrow(data)){
    if(is.na(data$steps[i])){
        NewData$steps[i]<- avrgPerInterval[(i %% 288)+1]
    }
}
```

1. Plot the new histogamme.
2. Calculate new mean and medaine after removing missing values.
```{r}
totalStepsPerDay1 <- sapply(split(NewData$steps,NewData$date),sum,na.rm=T)
hist(totalStepsPerDay1)

meanPerDay1 <- mean(totalStepsPerDay1)
MedianPerDay1 <- median(totalStepsPerDay1)

```

The new mean of number of steps is **`r meanPerDay1`** and the old one is **`r meanPerDay`**.  
The new mediane is **`r MedianPerDay1`** and the old one is **`r MedianPerDay`**.  

## Are there differences in activity patterns between weekdays and weekends?

Function for detrmine using Date if this day is weekend or weekday.
```{r}
library("reshape")

weekEndDay<-function(x){
    if(weekdays(x)  %in% c( "Saturday","Sunday")){
        "Weekend"
    }
    else{
        "weekday"
    }      
}
```

1. Add new column in data idicate if this day is weekend or week day.
2. Cast data to detrmine in each weekend or week day in specific interval the number of step.
3. Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval       (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 



```{r}
weekend <- sapply(data$date,weekEndDay)
data <- cbind(data,weekend)

castedData <- cast(data,weekend~interval,value="steps",sum,na.rm=T)

par(mfrow=c(2,1))
plot(x=intervals,y=castedData[2,2:ncol(castedData)],type="l",xlab="interval",ylab="number of step",main="Weekend",col="blue")
plot(x=intervals,y=castedData[1,2:ncol(castedData)],type="l",xlab="interval",ylab="number of step",main="Weekday",col="blue")

```

#### Comparaison using standard deviation.
The standard deviation of steps in weekend **`r sd(castedData[2,2:ncol(castedData)])`** and in weekday **`r sd(castedData[1,2:ncol(castedData)])`**.  
#### Comparaison using standard Average.
The average of steps in weekend **`r mean(as.numeric(castedData[2,2:ncol(castedData)]))`**. and in weekday **`r mean(as.numeric(castedData[1,2:ncol(castedData)]))`**.  
#### Comparaison using maximum number of steps.
The maximum number of steps in weekend **`r max(as.numeric(castedData[2,2:ncol(castedData)]))`**. and in weekday **`r max(as.numeric(castedData[1,2:ncol(castedData)]))`**. 
